Project Thesis/Report

### 2022-01-10 Gandhi_Project_Report/draft3.tex
Done with S 4.6

\usepackage{enumitem}

@@

A core concept in the study of dynamical systems is that of invariance. 
\begin{Definition}
  [\bf Invariant Set] \label{Dfn_Flow}
  Given a discrete-time dynamical system $T: U \to U$, a subset $A \subset U$ is said to be an \emph{invariant set} if $T(A) =A$. 
\end{Definition}

We also define the \emph{orbit of $T$} to be a sequence $\bar{u} = \{u_n\}_{n\in \mathbb{Z}}$ obeying the update equation, $u_{n+1}=Tu_n$, $n \in \mathbb{Z}$.

@@

There are several intricately different definitions of chaos available in the literature \textbf{(Cite)}, with each of them indicating some aspect of complexity. \ednote%{B:  what are we saying here? M: Devaney's definition is not the only one, there are about half a dozen definitions as well}  
In practice, it is not possible to verify which definition or notion does the underlying dynamical system satisfies when only data is observed from it. We do, however, specifically recall the definition of chaos in the sense of Devaney \cite{devaney2018introduction,de2013elements} so as to understand some nuances behind the complexity. Devaney's definition of chaos has three requirements, with the first being the notion  of sensitive dependence on initial conditions.

@@

\begin{Definition}
  [\bf Nonautonomous Dynamical System (adopted from~\cite{Manju_ESP}))] \label{Dfn_Soln}\rm
  A nonautonomous dynamical system (NDS) on a space $X$ is simply a dynamical system comprising a family of maps $\{g_n\}_{n \in \mathbb{Z}}$ where $g_n(\ldotp):=g(u_n,\ldotp):X\to{X}$ is continuous. Each $g_n$ arises as a consequence of an input $u_n$ from the input space $U$, a topological space.  \ednote{Does the reader immediately remember/know what is meant by a dynamical system?}
\end{Definition}

@@

\begin{Example}\label{Ex_exp} \rm 
 Consider the driven system $g(u_n,x_n)=x^n$ for $X=[0,1]$, $U=\mathbb{R}$. The system has an uncountable number of solutions, as there exists a solution for every $x\in{X}$ which also passes through the point $x$ and $lim_{n\to\infty}x_n=0$, $lim_{n\to{-}\infty}x_n=1$. 
 The proof is deferred to immediately after the next paragraph (see~\ref{rem_proofEx}).

@@

\begin{Definition}
  [\bf Reachable Set]\label{Dfn_ReachableSet}\rm
The reachable set of a driven system $g$ is exactly the union of all the elements of all the solutions, i.e., 
\[X_U :=\Big \{x \in X:  x = x_k \mbox{ where  $\{x_n\}$  is a solution for some  $\bar{u}$} \Big \}.\]
The set of all reachable states at a specific time $n$ for input $\overline{u}$ is denoted by $X_n(\overline{u})$
\end{Definition}

We note that $x\in{X_n(\overline{u})}$ if and only $g$ has a solution $\{x_k\}$ for $x_n=x$ and input $\overline{u}$. \textbf{(Cite)}.

@@

This will lead to a result established later, but which is worth taking note of now: \textit{$g$ being a topological contraction is equivalent to the existence of a unique entire solution.} To this end we define the notion of a topological contraction. \ednote{M: I do not think you need the definition of a topological contraction unless it is defined through a process. Please note that you have defined $X_n(\bar{u})$ differently than what was in the earlier works. Now it is just the collection of solution values at time $n$ with the input $\bar{u}$ and topological contraction is just paraphrasing the USP definition. B: Is it better to then just leave out the entire sentence?}

\begin{Definition}
  [\bf Topological Contraction]\label{Dfn_TopContr}\rm
  A function $g:U\times{X}\to{X}$ is a topological contraction if for all $n\in\mathbb{Z}$ and all $\overline{u}\subset{U}$, $X_n(\overline{u})$ is a singleton subset of $X$.  
\end{Definition}

\begin{Remark}
  [\bf Proof of Example~\ref{Ex_exp}] \label{rem_proofEx} \rm
  To see this, we show that for every input $\overline{u}=\{u_n\}_{n\in\mathbb{Z}}$, $g(u_n,\cdot)$ is a contraction map on $X$ for every $k\in\mathbb{Z}$. 
  Indeed $g(u_n,x_n)=x^n$ depends only on the state at time $n$ and so for each input $\overline{u}$, $X_n(\overline{u})=\{x^n\}$ is a singleton subset.
  We then easily conclude that $\{x_n\}\equiv0$ is the only solution. 
\end{Remark}

@@

\begin{Theorem}\label{Thm_CompactExistence}
 Let $g$ be a driven system.  If $X$ is compact then for each input $\overline{u}$, there exists at least one solution to the driven system $g(\ldotp, x)$
\end{Theorem}
\begin{proof}
  \textbf{I've muddled myself here.}
  % Consider an input $\overline{u}={\{u_n\}}_{n\in\mathbb{Z}}\subset{U}$ and driven system $g:U\times{X}\to{X}$  generating a sequence ${\{x_n\}}_{n\in\mathbb{Z}}$ in the compact space $X$ \ednote{This statement is not correct.}. 
  % Since $X$ is a metric space, it follows immediately from a well-known result in Analysis (\textbf{Cite}) that $\{x_n\}$ has a convergent subsequence which then is a solution.  \ednote{This is not a valid proof. Please refer to the proof existence of a solution in \cite{Manju_ESP}.}
\end{proof}

@@

------
xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx

@@

\begin{Theorem}
  For a compactly driven system, a causal mapping $H$ exists if and only if $g$ has the USP. 
\end{Theorem}
\begin{proof}
  \emph{See proof from p.15 of \cite{manjunath2021universal}}
  \textbf{Expand.}
\end{proof}

\textbf{When considering the map $H$, we once again restrict restrict our focus of $H$ to the function $H_2(\overline{u}):=(h(r\overline{u}, h(\overline{u})))$. }

@@

Before formalising this, we make mention of the concept of the inverse-limit system of a dynamical system $(U,T)$. 
  In broad sweeps, the inverse-limit system of a dynamical system is a self-map on a subset of an infinite-dimensional space \textbf{(Expand)} where each point in the inverse-limit space corresponds to a backward orbit of the map $T$. Following \cite{manjunath2021universal}, the inverse-limit space is denoted by $\widehat{U}_T\subset \overleftarrow{U}$, and defined as $\widehat{U}_T\subset: = \{ (\ldots,u_{-2},u_{-1}): Tu_{i} = u_{i+1}$. 
  The map $T$, then, induces a self-map $\widehat{T}: (\ldots,u_{-2},u_{-1}) \mapsto  (\ldots,u_{-2},u_{-1},T(u_{-1}))$ on $\widehat{U}_T$.

@@

\begin{Theorem}
 If $g(\ldotp,x)$ is invertible and has the USP then $H$ is a causal embedding. 
\end{Theorem}
\begin{proof}
  \emph{See proof from \cite{manjunath2021universal}}  \textbf{Expand.}
\end{proof}

@@

The next result is of relative import but to fully comprehend its influence, some discussion still remains. We state it in part and postpone the proof.
% Before stating it, we define a concept which will directly be shown to be equivalent to the USP.

\begin{Theorem}
  The following statements are equivalent:
  \vspace{-8mm}
  \begin{enumerate}[noitemsep, label=\roman*.]
    \item $g$ has the USP. 
    \item $g$ is a topological contraction.
  \end{enumerate}
\end{Theorem}
\begin{proof}
  (postponed) 
\end{proof}

@@

% \textbf{The proof of the result for the UAP is left to literature. Why? Cite.}

@@

It is worth taking note of a specific driven system. The function 
\begin{equation}  \label{eqn_driving}
  g(u,x) = (1-a)x + a\overline{\tanh}(Au + \alpha Bx)
\end{equation} 
is both SI-invertible and possesses the USP. 
One may easily show that $g$ is SI-invertible by recovering $u_n$ by 
\begin{equation} \label{eqn_SI_RNN}
  u_n := A^{-1}\bigg(\overline{\tanh}^{^{-1}}\frac{1}{a}\Big(x_{n+1}-(1-a)x_n\Big) \bigg) - \alpha B x_n
  \end{equation}
  for $x_{n-1}, x_n$ known.

The proof of these $g$ possessing the USP is rather involved and hence the proof, instead of being replicated here, may be found in \textbf{(cite relevant article)}
This specific driving function $g$ is used in our implementation and is more completely discussed in chapter \ref{sect5}

@@

As the Unique Attraction Property guarantees that all trajectories \ednote{(B: which ones?)} will converge to the same trajectory as time moves forward. 
Incredibly, this permits one to initialise a driven system $g$ with an altogether arbitrary initial value $y_m\in{X}$ where $m\in\mathbb{Z}$ and the UAP then guarantees that the sequence $\{y_{m+1}, y_{m+2},\ldots\}$ which satisfies the relation $y_n=g(u_n, y_{n-1})$  for  $k\geq{m}$  will uniformly approach the elements $\{x_n\}$ of the actual solution (See \cite[Th.11,]{Manju_Nonlinearity} or \cite[Eqn.18]{Supp}).

@@

\begin{Theorem}
  The following statements are equivalent:
  \vspace{-8mm}
  \begin{enumerate}[noitemsep, label=\roman*.]
    \item $g$ has the USP.
    \item $g$ is a topological contraction.
    \item $g$ has the UAP.
  \end{enumerate}
\end{Theorem}
\begin{proof}
  (We'll show this)  [11] eqn.6 
\end{proof}

@@

This also solves the problem of perturbation/noise introduced by the observable or measurement function. Recall from before that an observable is inherently a map that discretises the underlying continuous-time system. Moreover, measurement and device error introduce mistakes. Since we are working with a chaotic system displaying sensitive dependence on initial conditions, these small errors could potentially send the trajectory into a completely different attractor. 
In the presence of the UAP, however, small measurement errors introduced into the system do not pose the same danger as before since the trajectories will eventually tend towards the actual solution.
\ednote{(B: Complete enough?)}
\textbf{From the perspective of perturbation of an autonomous
dynamical system, if the resulting nonautonomous system has the USP, then the
temporal structure of noise relates to the dynamics on a nonautonomous uniform
attractor through a topological semi-conjugacy or a conjugacy.}

@@

\begin{Theorem}
    $(Y_T, G_T)$ is semi-conjugate to $(\widehat{U}, \widehat{T})$.
\end{Theorem}
\begin{proof}
  \textbf{Show}
\end{proof}

@@

@@

In \cite{manjunath2021universal} it is illustrated in an empirical fashion in that the map $G_T$ describes dynamics which are less functionally complex than that of $T$ or of the map $\Phi_{2d,\theta}$. This is done by implementing a Recurrent Neural Network (RNN), but not discussed as of yet. \ednote{(B: Add in that we'll be using Pearson coefficient?)}

@@

\begin{Theorem}
  When $x_n$ and $x_{n-1}$ are successive points on a solution obtained for input-orbit of $T$ $\{u_n\}$, then the map $\Gamma: X\times{X}\to{U}$ defined by $(x_{n-1},x_n)\mapsto{u_n}$ exists whenever $G_T$ exists 
\end{Theorem}
\begin{proof}
  The proof may be found in~\cite[Th. 3c]{manjunath2021universal}.
\end{proof}

@@

\begin{figure}[ht] \label{Fig_Gamma}
  \includegraphics[scale=0.3]{_gammatemp.png}
  \centering
  \label{fig:conjugacy_with_Gamma}
\end{figure}

@@

Learning $\Gamma$ saves computational resources. This is due to the fact that the input $u_n$ lies in a lower-dimensional subspace of the high-dimensional space $X\times{X}$. In practice, if the input is of a  lower dimension, one may easily embed it into the space $X\times{X}$ by padding the vector $u_n$ with zeroes.

\textbf{Can we discuss how to put this across well? I would like to add it in as I think it beneficial to the discussion.}

Moreover, two types of stability are achieved when $g$ has the USP.

Input- and parameter-related stability is achieved if $g$ has the USP
\newline \emph{Define/Explain the two types of stabilities}
\newline \emph{Sentence on $u$, $v$, tails, USP and product topology}
\newline If we opt to work with $Y_T$, errors can occur. Expand
\newline $\Gamma$ is more stable. Explain why.

Thirdly, $\Gamma$ is globally dissipative whereas $G_T$ is not guaranteed to be(\emph{Cite}). \ednote{B: Add linking sentence.} Global dissipativity prevents large numerical errors due to input data.

The $\omega$-limit set $\omega(u;T)$ of a point $u$ is defined to be the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$. This set is nonempty, and $\omega(u;T)$ is invariant. \textbf{(Change)}

\begin{Definition}
  [\bf Global Dissipativity]\label{Dfn_GlobalDiss}\rm
	We say a dynamical system $(Z,f)$ is globally dissipative if there exists a nonempty proper closed subset $B$ of $Z$  so that for all $x\in Z\setminus B$, (i). $\omega(x;f) \subset B$  and (ii).  $B$ is positively invariant that is $f(B) \subset B$. We call any such closed subset $B$ a trapping set of $f$. 
\end{Definition}

If a system is not globally dissipative, neglibly small errors result in major errors which thus induce predictions to fail utterly. If one learns $G_T$ directly, we are in danger of replicating the problems arising for Takens.

@@

In this section we turn to the implementation of the function $G_T$ by discussing a method to learn the function $\Gamma$. We then indicate where this implementation differs from previous work done by (\cite{manjunath2021universal}) and show that our results do hold water by considering the Pearson correlation coefficient. \ednote{(this sentence needs to be changed) Are you talking of MDD? Or the sparse matrix?.} 
Finally, we present and discuss some numerical results for a focus system of this project: the double pendulum. We also consider other chaotic attractors.

\section{Implementing $G_T$}
The map $G_T$ is implemented by applying the functions $\Gamma$, $g$ and the projection function $\pi_2$ in the relevant order. To reach this point, one must first learn $\Gamma$.
After 1.normalising the data to ensure it lies within the interval $|x| < 1$ and has zero mean value, we 2.drive the system with external input through the function $g$.

From before (\eqref{eqn_driving}): $ g(u,x)=(1-a)x+a{\cdot}\tanh(Au+\alpha{a}Bx)$. Remember that $g$ is SI-invertible. $A$ is the so-called “input matrix” and $B$ the “reservoir matrix”.

We make use of a Recurrent Neural Network (RNN) in our implementation. RNN's are widely used in the literature for discrete-time processing and are considered to be discrete-time version of continuous dynamical systems modelling a network of neurons. (\textbf{Cite.})
A complete discussion of the RNN is outside the scope of this thesis, so we only present the details necessary to the current discussion. For information, (\textbf{Cite, Cite, Cite}) may be consulted.

If the RNN we will be using in the next step has $N$ neurons, then $A$ and $B$ are each $N\times{N}$ matrices. Inputs $u_n$ of dimension $K < N$ are padded by zeroes to obtain an input of dimension $N$.

No need for entire left-infinite history of input: An arbitrary initialisation of the system with initial values and the driving of the system with the function $g$ means that the generated values will come to approximate the actual solution’s elements in a uniform manner thanks to the UAP.

We then 3.set up/prepare the RNN. Utilising a network $N$ neurons (the same $N$ as in the dimensions of the matrices $A,B$). Training is accomplished using the Adam Optimiser where the MSE loss function is optimised. The system is trained several times(in our case 4) with incrementally smaller learning rates on each run (0.001 and thereafter divided by 10 at each run). 
\ednote{Please present the number of layers in the network, number of neurons in each layer and the length of the training validation and testing for learning Gamma.}

@@

The two model equations for learning from data (\eqref{eqns_from_data}) also allows us to determine what the state $x_{n+1}$ at time $n+1$ will be. 
We specifically remind ourselves of \[x_{k+1}=\pi_2 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k)\]

@@

\section{Numerical simulations}
\subsection{Double Pendulum}
\begin{itemize}
  \item What it is. 
  \item Why it's chaotic
  \item How data was generated for this. (noiseless)
  \item Why data dying down doesn't work (ie. Not surjective)
  \item Other possibilities
\end{itemize}

@@

An ideal system is assumed where the rods are massless and no friction in the pivots are considered.
One is easily reminded of the Lorenz system, another three-equation deterministic model which exhibits chaotic behaviour.
The scalar equations of motion are presented below and were derived with the use of Newtonian physics and can be found in \textbf{(Cite.)}

>>>>

\usepackage{enumitem}
\usepackage{multirow}

@@

A core concept in the study of dynamical systems is that of invariance. 
\begin{Definition}
  [\bf Invariant Set] \label{Dfn_InvariantSet}
  Given a discrete-time dynamical system $T: U \to U$, a subset $A \subset U$ is said to be an \emph{invariant set} if $T(A) =A$. 
\end{Definition}

We also define the orbit of a function $T$.
\begin{Definition}
  [\bf Orbit of $T$] \label{Dfn_Orbit}
  The orbit of $T$ is to be defined the sequence $\bar{u} = \{u_n\}_{n\in \mathbb{Z}}$ obeying the update equation, $u_{n+1}=Tu_n$, $n \in \mathbb{Z}$. 
\end{Definition}

@@

There are several intricately different definitions of chaos available in the literature \textbf{(Cite)}, with each of them indicating some aspect of complexity. %\ednote{B:  what are we saying here? M: Devaney's definition is not the only one, there are about half a dozen definitions as well}  
In practice, it is not possible to verify which definition or notion does the underlying dynamical system satisfies when only data is observed from it. We do, however, specifically recall the definition of chaos in the sense of Devaney \cite{devaney2018introduction,de2013elements} so as to understand some nuances behind the complexity. Devaney's definition of chaos has three requirements, with the first being the notion  of sensitive dependence on initial conditions.

@@

\begin{Definition}
  [\bf Nonautonomous Dynamical System (adopted from~\cite{Manju_ESP}))] \label{Dfn_NDS}\rm
  A nonautonomous dynamical system (NDS) on a space $X$ is simply a dynamical system comprising a family of maps $\{g_n\}_{n \in \mathbb{Z}}$ where $g_n(\ldotp):=g(u_n,\ldotp):X\to{X}$ is continuous. Each $g_n$ arises as a consequence of an input $u_n$ from the input space $U$, a topological space. 
\end{Definition}

@@

\begin{Example}\label{Ex_exp} \rm 
 Consider the driven system $g(u_n,x_n)=x^n$ for $X=[0,1]$, $U=\mathbb{R}$. The system has an uncountable number of solutions, as there exists a solution for every $x\in{X}$ which also passes through the point $x$ and $lim_{n\to\infty}x_n=0$, $lim_{n\to{-}\infty}x_n=1$. 
 The proof is deferred to immediately after the next paragraph (see~\ref{rem_proofEx}).
\end{Example}

@@

\begin{Definition}
  [\bf Reachable Set]\label{Dfn_ReachableSet}\rm
The reachable set of a driven system $g$ is exactly the union of all the elements of all the (entire) solutions, i.e., 
\[X_U :=\Big \{x \in X:  x = x_k \mbox{ where  $\{x_n\}$  is a solution for some  $\bar{u}$} \Big \}.\]
The set of all reachable states at a specific time $n$ for input $\overline{u}$ is denoted by $X_n(\overline{u})$
\end{Definition}

The reachable set itself can defined independently from the fact that $g$ does or does not have the ESP/USP. We note that $x\in{X_n(\overline{u})}$ if and only $g$ has a solution $\{x_k\}$ for $x_n=x$ and input $\overline{u}$. \textbf{(Cite)}.

@@

This will lead to a result established later, but which is worth taking note of now: \textit{$g$ being a topological contraction is equivalent to the existence of a unique entire solution.}

% \begin{Definition}
%   [\bf Topological Contraction]\label{Dfn_TopContr}\rm
%   A function $g:U\times{X}\to{X}$ is a topological contraction if for all $n\in\mathbb{Z}$ and all $\overline{u}\subset{U}$, $X_n(\overline{u})$ is a singleton subset of $X$.  
% \end{Definition}

\begin{Remark}
  [\bf Proof of Example~\ref{Ex_exp}] \label{rem_proofEx} \rm
  To see this, we show that for every input $\overline{u}=\{u_n\}_{n\in\mathbb{Z}}$, $g(u_n,\cdot)$ is a contraction map on $X$ for every $k\in\mathbb{Z}$. 
  Indeed $g(u_n,x_n)=x^n$ depends only on the state at time $n$ and so for each input $\overline{u}$, $X_n(\overline{u})=\{x^n\}$ is a singleton subset.
  We then easily conclude that $\{x_n\}\equiv0$ is the only solution. 
\end{Remark} \ednote{B: Is this sufficient?}

@@

\begin{Theorem}\label{Thm_CompactExistence}
 Let $g$ be a driven system.  If $X$ is compact then for each input $\overline{u}$, there exists at least one solution to the driven system $g(\ldotp, x)$
\end{Theorem}
\begin{proof}
  May be found in \cite{kloeden2011nonautonomous, manjunath2014dynamics, manjunath2013echo}
\end{proof}

@@

Historically, data have been transformed into a different space for processing to gain some advantages. For example, according to Vapnik–Chervonenkis theory (e.g., [16]) mapping data into a higher dimension often provides much greater classification power. 
Mapping temporal data through a delay-coordinate mapping in Takens delay embedding uncovers hidden information [17]. In data-driven modeling employing Koopman’s theory (e.g., [18]–[20]), 
approximating the dynamics of a dynamical system through a linear dynamicalsystem in a higher dimensional space is often possible by mapping the data samplewise into the higher dimensional space

we first note that when T is not invertible, the dynamics of (U, T ) is only topologically semi-conjugate to $(\hat{U}, \hat{T})$ and hence the dynamics in $(\hat{U}, \hat{T})$ could have additional complexity.

xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx

@@

\begin{Theorem}
  For a compactly driven system, a causal mapping $H$ exists if and only if $g$ has the USP. 
\end{Theorem}
\begin{proof}
  See \cite[Th.3]{manjunath2013echo}.
\end{proof}

\textbf{When considering the map $H$, we once again restrict our focus of $H$ to the function $H_2(\overline{u}):=(h(r\overline{u}, h(\overline{u})))$. }

@@

Before formalising this, we make mention of the concept of the inverse-limit system of a dynamical system $(U,T)$. 
These spaces are noteworthy as their topology can help one ascertain much about the original system \cite{Manju_IEEE}. The topological structure for inverse-limit systems of chaotic dynamical systems is complicated, so we relegate further discussion to \cite{kennedy2008inverse_limit, ingram2011inverse}.
In broad sweeps, the inverse-limit space of a dynamical system is a self-map on a subset of an infinite-dimensional space where each point in the inverse-limit space corresponds to a backward orbit of the map $T$. 
Following \cite{manjunath2021universal, ingram2011inverse}:

\begin{Definition}
  [\bf Inverse-Limit Space, Inverse-Limit System]\label{Dfn_InverseL}\rm
 The subset $\widehat{U}_T\subset \overleftarrow{U}$ defined by \[\widehat{U}_T: = \{ (\ldots,u_{-2},u_{-1}): Tu_{i} = u_{i+1}\]is called the inverse-limit space of $(U,T)$.
 The self-map $\widehat{T}$  induced by $T$ on $\widehat{U}_T$  is defined by \[\widehat{T}: (\ldots,u_{-2},u_{-1}) \mapsto  (\ldots,u_{-2},u_{-1},T(u_{-1}))\] and the resulting dynamical system $(\widehat{U_T}, \widehat{T})$ is called the inverse-limit system of $(U,T)$.
\end{Definition}

@@

\begin{Theorem}
 If $g(\ldotp,x)$ is invertible and has the USP then $H$ is a causal embedding. 
\end{Theorem}
\begin{proof}
  See \cite{manjunath2021universal}.
\end{proof}

@@

The next result is of relative import but to fully comprehend its influence, some discussion still remains. We state it in part here.
% Before stating it, we define a concept which will directly be shown to be equivalent to the USP.

\begin{Theorem}
  The following statements are equivalent:
  \vspace{-8mm}
  \begin{enumerate}[noitemsep, label=\roman*.]
    \item $g$ has the USP. 
    \item $g$ is a topological contraction.
  \end{enumerate}
\end{Theorem}
% \begin{proof}
%   (postponed) 
% \end{proof}

@@

@@

It is worth at this stage taking note of a specific driven system in the form of a discrete state-space model that has acquired some adherence in applications \cite{Manju_IEEE} and especially those pertaining to Echo State Networks and ESP. The function

\begin{equation}  \label{eqn_driving}
  g(u,x) = (1-a)x + a\overline{\tanh}(Au + \alpha Bx)
\end{equation} 
is both SI-invertible and, if $\alpha\beta<1$, also possesses the USP. 
One may easily show that $g$ is SI-invertible by recovering $u_n$ by 
\begin{equation} \label{eqn_SI_RNN}
  u_n := A^{-1}\bigg(\overline{\tanh}^{^{-1}}\frac{1}{a}\Big(x_{n+1}-(1-a)x_n\Big) \bigg) - \alpha B x_n
  \end{equation}
  for $x_{n-1}, x_n$ known.

The proof that the map $g$ possesses the USP is rather involved and hence, instead of being replicated here, may be found in \cite[Th.2]{manjunath2013echo }
This specific driving function $g$ is used in our implementation and is discussed more completely in chapter \ref{sect5}

@@

As the Unique Attraction Property guarantees that all trajectories will converge to the same trajectory as time moves forward. 
Incredibly, this permits one to initialise a driven system $g$ with an altogether arbitrary initial value $y_m\in{X}$ where $m\in\mathbb{Z}$ and the UAP then guarantees that the sequence $\{y_{m+1}, y_{m+2},\ldots\}$ which satisfies the relation $y_n=g(u_n, y_{n-1})$  for  $k\geq{m}$  will uniformly approach the elements $\{x_n\}$ of the actual solution (See \cite[Th.11,]{Manju_Nonlinearity} or \cite[Eqn.18]{Supp}).

@@

\begin{Theorem}\label{thm_usp-contr-uap}
  The following statements are equivalent:
  \vspace{-8mm}
  \begin{enumerate}[noitemsep, label=\roman*.]
    \item $g$ has the USP.
    \item $g$ is a topological contraction.
    \item $g$ has the UAP.
  \end{enumerate}
\end{Theorem}
\begin{proof}
  May be found in \cite[Th.1]{Manju_Nonlinearity}
\end{proof}

@@

This also solves the problem of perturbation/noise introduced by the observable or measurement function. Recall from before that an observable is inherently a map that discretises the underlying continuous-time system. Moreover, measurement and device error introduce mistakes. Since we are working with a chaotic system displaying sensitive dependence on initial conditions, these small errors could potentially send the trajectory into a completely different attractor. 
In the presence of the UAP, however, small measurement errors introduced into the system do not pose the same danger as before since the trajectories will eventually tend towards the actual solution.
\ednote{(B: Complete enough?)}
% \textbf{From the perspective of perturbation of an autonomous dynamical system, if the resulting nonautonomous system has the USP, then the temporal structure of noise relates to the dynamics on a nonautonomous uniform attractor through a topological semi-conjugacy or a conjugacy.}

@@

\begin{Theorem}
    $(Y_T, G_T)$ is semi-conjugate to $(\widehat{U}, \widehat{T})$.
\end{Theorem}
\begin{proof}
  See \cite[Th.3, Th.4]{manunath2021universal}
\end{proof}

@@

% Below is reproduced a pictorial representation of the relationship between different entities and properties. \textbf{(Cite SIG_25May)}
% Below is reproduced a pictorial representation of the relationship between the different entities and their properties.

\begin{figure}[ht]
  \includegraphics[scale=0.3]{_summarypictorial.png}
  \centering
  \label{fig:fig_pictorialSummary}
\end{figure}

@@

In \cite{manjunath2021universal} it is illustrated in an empirical fashion in that the map $G_T$ describes dynamics which are less functionally complex than that of $T$ or of the map $\Phi_{2d,\theta}$. This is done by implementing a Recurrent Neural Network (RNN), which is discussed in~\ref{ch5}. \ednote{(B: Should I add in that we'll be using Pearson coefficient?)}

@@

\begin{Theorem}
  When $x_n$ and $x_{n-1}$ are successive points on a solution obtained for input-orbit of $T$ $\{u_n\}$, then the map $\Gamma: X\times{X}\to{U}$ defined by $(x_{n-1},x_n)\mapsto{u_n}$ exists whenever $G_T$ exists 
\end{Theorem}
\begin{proof}
  See \cite[Th. 3c]{manjunath2021universal}.
\end{proof}

@@

\begin{figure}[ht]
  \includegraphics[scale=0.3]{_gammatemp.png}
  \centering
  \label{fig:temp_Gamma}
\end{figure}

@@

In the first place, Learning $\Gamma$ saves computational resources. This is due to the fact that the input $u_n$ lies in a lower-dimensional subspace of the high-dimensional space $X\times{X}$ and we can learn $\Gamma$ on a small subset of the space. In practice, if the input is of a  lower dimension, one may easily embed it into the space $X\times{X}$ by padding the vector $u_n$ with zeroes.

Secondly, the function $\Gamma$ is known to be stable. Learning Gamma makes use of equations \ref{eqns_from_data} which in turn employs $g$, a function possessing the USP, and offers distinct advantages with regards to stability in the presence of perturbation. It is desirable that inputs that differ only slightly would have only ‘slight’ effects on the output of the system. Moreover, one would hope to prevent numerical errors originating from input and measurement noise (see~\cite[Th. 5]{manjunath2021universal}). 
As the data fed into the system is a function of both the input given and the parameters of the system itself (i.e. the value a, $\alpha$ in $\tanh$), we investigate both input-related and parameter-related stability. Both these notions are defined by way of the continuity of an encoding map of an input and design parameter respectively in \cite{manjunath2020stability}.\

Informally, we may consider input-related stability as concerned with the question of whether or not small variations in input would result in small responses.   
If two inputs $u$ and $u_noisy$ are close in the product topology, then their tails could differ greatly due to the metric's insensitivity to the differences in sequence's tails.
 However if the system has the USP then the function $h$ is continuous and so $h(u), h(u_noisy)$ remain near – maintaining a measure of robustness to input and measurement noise.\cite{manjunath2021universal}.
In a similar fashion, parameter-related stability is obtained thanks to a result \cite[Lemma 3.2]{manjunath2020stability} relating this form of stability with the ESP (which is here equivalent to the USP).

% Secondly,  due to $g$ possessing the USP, both input- and parameter-related stability~\cite{manjunath2020stability} is obtained, which in turn prevents numerical errors that caused by input and measurement noise (see~\cite[Th. 5]{manjunath2021universal}).

\begin{figure}[ht]
  \includegraphics[scale=0.4]{_AutovsDriven.eps}
  \centering
  \label{fig:learningFailure}
  \captionof{figure}{Illustrating empirically the importance of global dissipativity. Plotted below the three principal components of randomly initiated trajectories of a learnt version of $G_T$ versus the trajectory $(x_{n-1}, x_n)$(grey) of driven stated obtained from the Thomas Attractor's data (see below). }
\end{figure}

\ednote{B: Would you please help me with the writing here? Otherwise I shall leave this out.}
\emph{Thirdly, $\Gamma$ is globally dissipative whereas $G_T$ is not guaranteed to be~\cite{manjunath2021universal}). 
Global dissipativity prevents large numerical errors due to input data.If a system is not globally dissipative, neglibly small errors result in major errors which thus induce predictions to fail utterly. If one learns $G_T$ directly, we are in danger of replicating the problems arising for Takens. }

% The $\omega$-limit set $\omega(u;T)$ of a point $u$ is defined to be the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$. This set is nonempty, and $\omega(u;T)$ is invariant. \textbf{(Change)}
% \begin{Definition}
%   [\bf Global Dissipativity]\label{Dfn_GlobalDiss}\rm
% 	We say a dynamical system $(Z,f)$ is globally dissipative if there exists a nonempty proper closed subset $B$ of $Z$  so that for all $x\in Z\setminus B$, (i). $\omega(x;f) \subset B$  and (ii).  $B$ is positively invariant that is $f(B) \subset B$. We call any such closed subset $B$ a trapping set of $f$. 
% \end{Definition}

@@

In this section we turn to the implementation of the function $G_T$ by discussing a method to learn the function $\Gamma$. We then indicate where this implementation differs from previous work done by (\cite{manjunath2021universal}) and empirically show that the SDD of $G_T$ are less functionally complex than $\Phi_{k, \theta}$ or $T$ by considering the Pearson correlation coefficient. 
Finally, we present and discuss some numerical results for a focus system of this project: the double pendulum. We also consider other chaotic attractors.

\section{Implementing $G_T$}
The map $G_T$ is implemented by applying the functions $\Gamma$, $g$ and the projection function $\pi_2$ in the relevant order as described by the equations~\ref{eqns_from_data}. To reach this point, one must first learn $\Gamma$.
After first normalising the data to ensure it lies within the interval $|x| < 1$ and has zero mean value, we continue by driving the system with external input through the function $g$.

From \eqref{eqn_driving} we know that $ g(u,x)=(1-a)x+a{\cdot}\tanh(Au+\alpha{a}Bx)$. Recall that $g$ is SI-invertible. %Adopting the terminology of \cite{manjunath2021universal}, $A$ is called the “input matrix” and $B$ the “reservoir matrix”.

We make use of a Recurrent Neural Network (RNN) in our implementation. RNN's are widely used in the literature for discrete-time processing and are considered to be discrete-time version of continuous dynamical systems modelling a network of neurons. (\textbf{Cite.})
A complete discussion of RNN's is outside the scope of this thesis, so we present only the details necessary to the current discussion. For information, (\textbf{Cite, Cite, Cite}) may be consulted. We implement our RNN in the \emph{Python} programming language and making use of the \emph{Scipy} and \emph{Tensorflow} libraries.

If the RNN we will be using in the next step has $N$ neurons, then $A$ and $B$ are each $N\times{N}$ matrices. Inputs $u_n$ of dimension $K < N$ are padded by zeroes to obtain an input of dimension $N$. As we do not need the entire left-infinite history of input, an arbitrary initialisation of the system with initial values followed by the driving of the system with the function $g$ means that the generated values will approach the actual solution’s elements in a uniform manner thanks to the UAP. \textbf{(Cite.)}

Thirdly we prepare the network. Utilising a network $N$ neurons (the same value $N$ as in the dimensions of the matrices $A,B$). Training is done using the Adam Optimiser where the MSE loss function is optimised. The system is trained several times(in our case 4) with incrementally smaller learning rates on each run (0.001 and thereafter divided by 10 at each run). 
Additional parameters such as the number of network-layers, neurons per layer and training-length are presented case-by-case below.

@@

The two model equations for learning from data (\eqref{eqns_from_data}) allows us to determine what the state $x_{n+1}$ at time $n+1$ will be. 
We specifically remind ourselves of \[x_{k+1}=\pi_2 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k)\]

@@

\section{Numerical simulations}
\subsection{Double Pendulum}

@@

This system has four variables: angle and angular velocity of each pendulum rod. An ideal system is assumed where the rods are massless and no friction in the pivots are considered.
The scalar equations of motion are derived using Newtonian physics and can be found in \cite{DPFormulas}. The second-order system is converted to a first-order ODE and solved using the Runge-Kutta numerical method.

This system is a classical example of a system exhibiting chaos. \textbf{Cite.}
